---
title: "Cluster summaries"
output: 
  html_document: 
    toc: yes
    code_folding: 'show'

runtime: shiny
---  


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

This report summarises an the results of an analysis to group English general practices using k-means clustering. Full details of the method can be found [here]("http://rpubs.com/jflowers/243520").

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(readr)
library(janitor)
library(purrr)
library(tidyr)
library(ggplot2)

gpdata <- read_csv("cluster.csv", col_types = c("ccddddddddii")) %>%
  clean_names() %>%
  rename(cluster1 = `_cluster`) %>%
  mutate(percent_eth = 100 * percent_eth)
```

```{r}
sliderInput("cluster", "Select cluster", min = 1, max = 15, value = 5)
```


## Cluster summary
`r renderText(paste0("Summary for Cluster ", input$cluster))`: mean values for the cluster.



```{r, echo = FALSE}
options(digits = 2)

renderTable(
gpdata %>% 
  filter(cluster1 == input$cluster) %>% 
  select(3:11) %>%
  map(mean) %>%
  map(round,2) %>%
  data.frame() %>%
  select(Ethnicity = percent_eth, 
         Practice_size = prac_pop, 
         Deprivation_score = imd2015,
         `Under_5%` = percent_5, 
         `Five_to_14%` = percent_5_14,
         `Under_18%` = percent_18,
         `Over_65%` = percent_65, 
         `Over_75%`= percent_75, 
         `Over_85%` = percent_85
         ) 
)
```

## Cluster profiles

This cluster contains `r renderText(paste0(gpdata %>% count(cluster1) %>% filter(cluster1 == input$cluster) %>% select(n), " practices"))`

### Interpreting the cluster profile plots

These plots show the z-score for each indicator for the cluster. The red dots show the z-score for each cluster and the dotted centre line at 0 is the national average value.

A z-score is a normalised score - if a score is 0 then this is the national average. A z-score of 1 is 1 standard deviation (SD) higher than the national average; a score of -1 is 1 SD lower and so on. 

For example if the `imd2015` is to the right of the centre line, this means on average practices in this cluster are more deprived than England as a whole. If `percent_eth` is to the left of the centre line, this means on average these practices have a lower than average proportion of people with recorded white ethnicity than England as a whole. These practice will tend to be more ehthnically diverse.

The higher the z-score for any indicator, the further from the national average practices in the relevant cluster will tend to be.

```{r, echo=FALSE}

## Scale dataset

gpdatas <- gpdata %>%
  select(3:11) %>% 
  map(scale) %>%
  data.frame() %>%
  ungroup()


renderPlot({
  gpdatas %>%
  cbind(gpdata[, 12]) %>%
  gather(indicator, value,1:9) %>%
  group_by(cluster1, indicator) %>%
  summarise(medval = median(value, na.rm = TRUE)) %>%
  filter(cluster1 == input$cluster) %>%
  ggplot(aes(indicator, medval)) +
  geom_point(colour = "red") +
  geom_line(aes(group = cluster1), colour = "blue")+
  geom_hline(yintercept = 0, lty = "dotted") +
  coord_flip()+
  labs(caption = "Black line = mean value", 
       title  ="Cluster profile plot", 
       subtitle = paste("Cluster: ", input$cluster), 
       y = "z-score") +
  theme(legend.position = "") +
  theme_bw() +
  ylim(c(-3, 3)) +
  geom_text(x = 9, y = -1.5, label = "below average", size = 2) +
  geom_text(x = 9, y = 1.5, label = "above average", size = 2)
}
)
```

### Practices in this cluster

The interactive table below shows the practices in `r renderText(paste0("Cluster ", input$cluster))`. 

```{r, echo=FALSE}
renderDataTable({ 
gpdata %>%
  filter(cluster1 == input$cluster) %>%
  select(Practice = x_area, CCG = ccg, cluster1) 
}
)

```

```{r, warning=FALSE, message = FALSE, fig.height= 8, results='asis', fig.align='center'}
# library(ggmap)
# 
# 
# gpmap <- read_tsv("http://data.gov.uk/data/resource/nhschoices/GP.csv", col_types = c("icccccccccccccddcccccc"))
# 
# gpmap <- gpmap %>% clean_names() %>%
#   select(organisationcode, parentname, latitude, longitude)
# 
# gpdata1 <- gpmap %>%
# left_join(gpdata, by = c("organisationcode" = "x_area")) %>%
#   filter(cluster1 == input$cluster)

# mymap1 <- get_map("England", source = "google", maptype = "roadmap",  zoom = 6)
# 
# map1 <- ggmap(mymap1) + 
#   geom_jitter(aes( y = latitude, x= longitude, colour = cluster1), 
#                            data = gpdata1, size = 1.5, alpha = 0.9) + 
#   coord_map(xlim = c(-6, 2.5), 
#             ylim = c(49.5, 55.5))
```

```{r, message=FALSE, warning=FALSE}
library(ggmap) 

 
 
gpmap <- read_tsv("http://data.gov.uk/data/resource/nhschoices/GP.csv", col_types = c("icccccccccccccddcccccc")) %>% clean_names()
gpdata1 <- gpmap %>%
  left_join(gpdata, by = c("organisationcode" = "x_area"))
```

## Map of practices in this cluster

```{r, message=FALSE, warning=FALSE}
renderPlot({
ggmap(get_map("England", source = "google", maptype = "roadmap",  zoom = 6)) +
  geom_jitter(aes( y = latitude, x= longitude, colour = cluster1), 
                           data = filter(gpdata1, cluster1 == input$cluster), size = 1.5, alpha = 0.5)+ coord_map() +
  labs(title = paste("Geographical distribution of practices in Cluster", input$cluster) ,  y = "", x = "") + 
  theme(axis.text = element_blank(), legend.position = "")
}
)
```


